---
output: html_document
editor_options: 
  chunk_output_type: console
---
```{r}
library(readxl)
library(xlsx)
library(ggplot2)
library(stringr)
library(ggbeeswarm)
library(dplyr)
library(ggpmisc)
library(Hmisc)
library(gtable)
library(ftExtra)
library(flextable)
library(grid)
library(cowplot)
library(svglite)
library(egg)
library(ggpubr)
library(dr4pl)
library(purrr)
library(gt)
library(tidyr)
library(VCA)
library(pracma)
```

```{r}
set.seed(123) ## to make sure geom_jitter() stays the same each run

rsq <- function(df) { 
  
  #https://stackoverflow.com/questions/40901445/function-to-calculate-r2-r-squared-in-r
  #https://en.wikipedia.org/wiki/Coefficient_of_determination#As_squared_correlation_coefficient
  
  obs <- df$data$Response
  res <- residuals(df)
  
  mean_obs <- mean(obs)
  SSres <- sum(res^2)
  SStot <- sum((obs-mean_obs)^2)
  
  1-(SSres/SStot)
}

RSV_A_CONV = 0.37

```

# Figure 1: ELISA, linearity, precision

## panel 1A: ELISA
```{r}
elisa_df = read_excel('paper_audit.xlsx', sheet = "Validation; ELISA")
elisa_df = elisa_df %>%
  group_by(`sample ID`) %>%
  mutate(ND50 = mean(ND50)) %>%
  mutate(co_norm = mean(co_norm)) %>%
  distinct(`sample ID`, ND50, .keep_all = TRUE)
  
elisa_df$ND50 = elisa_df$ND50 * RSV_A_CONV
elisa_df$logND50 = log2(elisa_df$ND50)

e_rubella = elisa_df %>%
  filter(str_detect(`sample ID`, "^R-"))

e_contrived_new = elisa_df %>%
  filter(str_detect(`sample ID`, "^lin #"))

#####################################
########## FITTING FORMULA ##########
#####################################
fit = dr4pl(data = e_contrived_new, 
            dose = ND50, 
            response = co_norm,
            method.init = "Mead",
            trend = "increasing",
            init.parm = dr4pl_theta(theta_1 = 200, theta_2 = 10000, theta_3 = 0.5, theta_4 = -1, isLog10 = FALSE),
            upperl = c(1000, 15000, 1.2, 100),
            lowerl = c(0, -Inf, 0.3, -100)
            )

fit_rq = rsq(fit)

parameters <- fit$parameters
pred <- tibble(Dose = 10 ^ seq(0.75, 4.6, 0.01)) %>%
  mutate(Response = MeanResponse(parameters, Dose))

pred = pred %>%
  filter(Dose < 6500.0)


my_form = function(x) coef(fit)[0] + (coef(fit)[3] - coef(fit[0]))/(1+(x/coef(fit)[1])^coef(fit)[2])

fig_elisa = ggplot(elisa_df, aes(x=ND50, y=co_norm)) + 
  
  
  scale_x_log10(limits = c(1, 10^5),
                breaks = c(10, 100, 1000, 10000, 100000),
                labels = scales::trans_format("log10", scales::math_format(10^.x)),) + 
  
  scale_y_continuous(limits = c(-10, 70), breaks = c(seq(0, 75, 10))) + 

  
  geom_line(data = pred, aes(x=Dose, y=Response), 
            color = 'black', 
            linetype = "dashed") + 
  
  geom_segment(x = -5, xend = 10000, y = 10, yend = 10, color = 'darkgrey') + 
  
  
  
  geom_point(data = elisa_df, size = 0.9, color = 'white') +
  geom_point(shape = 1, data = e_rubella, size = 0.9, color = 'blue', stroke = 0.6) +
  geom_point(shape = 1, data = e_contrived_new, size = 0.9, color = 'black', stroke = 0.6) +
  annotation_logticks(sides = 'b') + 
  theme_classic() + 
  annotate("text", x = 10, y = 65, label="R^2==0.99", parse = TRUE, size = 3) + 
  xlab('IU/mL estimated by RSV FRNT assay') + 
  ylab('Absorbance relative to ELISA cutoff control') + 
  theme(axis.text.x = element_text(size = 8, color = 'black', hjust = 0.4),
        axis.text.y = element_text(size = 8, color = 'black'),
        axis.title.x = element_text(size = 8), 
        axis.title.y = element_text(size = 8),
        plot.background = element_rect(fill = '#00000000'),
        panel.background = element_rect(fill = '#00000000'),
        panel.border = element_blank(),
        aspect.ratio = 1
        )

print(summary(fit))
fig_elisa

# below: paper audit version of curve fit
```

## panel 1B: linearity
```{r}
### panel B - linearity

lin_df = read_excel('paper_audit.xlsx', sheet = "Validation; ELISA") %>%
  filter(str_detect(`sample ID`, "^lin #")) %>%
  mutate(ND50 = ND50 * RSV_A_CONV)

formula = y ~ poly(x, 1, raw = TRUE)

myformat <- "y = %sx + %s"

################################
## get fit stats
#####################
# dat <- read.csv("Qualification/Pooled_qual_data.csv")
dat_l <- read_excel("paper_audit.xlsx", sheet = "Validation; ELISA") %>%
  mutate(cumm_dil = 1 / dil_factor, sample = `sample ID`)

dat_l_ND50 <- dat_l
dat_l_ND50$ND50 = dat_l_ND50$ND50 * 0.37


AMR_50 <- lm(log10(ND50)~log10(1/cumm_dil),dat_l_ND50)
AMR_50_sum <- summary(AMR_50)

row_names = c(
  "r-squared", 
  "adjusted r-squared", 
  "residual S.E.", 
  "f-statistic", 
  "p-value",
  "n",
  "slope",
  "intercept"
  )

## table for ND50 lm value
f_50 = AMR_50_sum$fstatistic

AMR_50_rows = c(
  round(AMR_50_sum$r.squared, 3),
  round(AMR_50_sum$adj.r.squared, 3),
  round(AMR_50_sum$sigma, 2),
  round(AMR_50_sum$fstatistic[1], 0),
  pf(f_50[1], f_50[2], f_50[3], lower.tail = FALSE),
  length(AMR_50$fitted.values),
  as.double(round(AMR_50$coefficients[2], 0)),
  round(AMR_50$coefficients[1], 2)
)

matrix_50 = as.data.frame(cbind(row_names, AMR_50_rows))
row.names(matrix_50) <- NULL
colnames(matrix_50) <- c("statistic", "value")

range_df <- dat_l_ND50 %>% 
  group_by(sample) %>%
  summarise(across(where(is.numeric), mean, na.rm = TRUE))

sprintf("Linearity: Mean IU/mL results ranging from %.*f to %.*f", 0, min(range_df$ND50), 0, max(range_df$ND50))

fig_linearity = ggplot(data = lin_df, aes(x=dil_factor, y=ND50)) + 
  

  stat_poly_line(data = lin_df, color = 'black', se = FALSE, size = 0.7, linetype = 'dashed'
                 ) + 
  
  geom_point(data = lin_df, size = 1.5, color = 'black', stroke = NA, fill = 'black', alpha = 0.6) + 
  
  stat_poly_eq(data = lin_df, 
               coef.digits = 3, 
               coef.keep.zeros = FALSE,
               color ='black', 
               size = 3,
               use_label("eq"),
               label.x = 0.2,
               label.y = 0.9) + 
  
  scale_y_log10(
    limits = c(15.5, 10^5),
    breaks = scales::trans_breaks("log10", function(x) 10^x),
    labels = scales::trans_format("log10", scales::math_format(10^.x)),
    ) + 
  
  scale_x_log10(
    limits = c(9^-3, 1),
    breaks = c(10^-3, 10^-2, 10^-1, 10^0),
    labels = scales::trans_format("log10", scales::math_format(10^.x)),
                ) + 
  
  annotation_logticks(sides = 'lb',
                      short = unit(0.075, "cm"),
                      mid = unit(0.15, "cm"),
                      long = unit(0.2, "cm")) + 
  theme_classic() + 
  xlab('Dilution factor') + 
  ylab("Neutralizing titer (IU/mL)") + 
  theme(
    axis.text.x = element_text(size = 8, color = 'black'),
    axis.text.y = element_text(size = 8, color = 'black'),
    plot.background = element_rect(fill = '#00000000'),
    panel.background = element_rect(fill = '#00000000'),
    panel.border = element_blank(),
    axis.title.x = element_text(size = 8),
    axis.title.y = element_text(size = 8),
    aspect.ratio = 1)
  
print(fit)
fig_linearity
```

## panel 1C: precision - linearity panel 
```{r}
cv_calc <- function(x) {
  data.frame(
    y = mean(x),
    ymin = mean(x) - 2 * sd(x),
    ymax = mean(x) + 2 * sd(x)
  )
}

test_calc <- function(x) {
  data.frame(
    y = 2^mean(log2(x)),
    ymin = 2^(mean(log2(x)) - 2 * log2(1.43)),
    ymax = 2^(mean(log2(x)) + 2 * log2(1.43))
  )
}

lprec_df = read_excel("paper_audit.xlsx", sheet = "Validation; Precision contrived")
lprec_df$ND50 = lprec_df$ND50 * RSV_A_CONV
lprec_df = lprec_df %>% mutate(
  tick = case_when(
    `sample ID` == "lin #2" ~ 2,
    `sample ID` == "lin #3" ~ 3,
    `sample ID` == "lin #4" ~ 4,
    `sample ID` == "lin #5" ~ 5
  )
)


lprec_df$logND50 = log2(lprec_df$ND50)

lprec_df = lprec_df %>%
  filter(plateID !="2023-Apr-25_plate5")

temp <- lprec_df %>%
  group_by(`sample ID`) %>%
  summarise(
    y = 2^mean(log2(ND50)),
    ymin = 2^(mean(log2(ND50)) - 2 * log2(1.43)),
    ymax = 2^(mean(log2(ND50)) + 2 * log2(1.43)),
    tick = mean(tick)
  )

sprintf("Expected IU/mL values ranging from %.*f to %.*f", 0, min(lprec_df$theoND50), 0, max(lprec_df$theoND50))

l_prec = ggplot(lprec_df, aes(x = tick, y = ND50, group = tick)) +
  
  annotate(
    geom = "rect",
    xmin = temp$tick-0.3,
    xmax = temp$tick+0.3,
    ymax = temp$ymax,
    ymin = temp$ymin,
    alpha = 0.3,
    fill = 'green'
  ) +
  
  annotate(
    geom = "segment", 
    x = temp$tick - 0.3,
    xend =temp$tick + 0.3,
    y = temp$y,
    yend = temp$y,
    linetype = 'solid',
    size = 0.15
  ) + 
  
  geom_jitter(size = 1.5, alpha = 0.6, color = 'black', stroke = NA,
             width = 0.1, 
             height = 0) + 
  
  scale_x_continuous(labels = c("2", "3", "4", "5"),
                     breaks = c(2, 3, 4, 5)) +
  
  scale_y_log10(
    labels = scales::trans_format("log10", scales::math_format(10^.x)),
    breaks = c(1, 10, 100, 1000, 10000, 10000),
    limits = c(1, 11^4),
    expand = c(0,0)
  ) +
  theme_classic() +
  xlab('Panel member') + 
  ylab('Measured IU/mL') + 
  annotation_logticks(sides = 'l') + 
  theme(
    axis.text.x = element_text(size = 8, color = 'black'),
    axis.text.y = element_text(size = 8, color = 'black'),
    axis.title.x = element_text(size = 8), 
    axis.title.y = element_text(size = 8),
    plot.background = element_rect(fill = '#00000000'),
    panel.background = element_rect(fill = '#00000000'),
    panel.border = element_blank(),
    aspect.ratio = 1
)

l_prec
```

## panel 1D: precision - rubella panel 
```{r}
cv_calc <- function(x) {
  data.frame(
    y = mean(x),
    ymin = mean(x) - 2 * sd(x),
    ymax = mean(x) + 2 * sd(x)
  )
}

test_calc <- function(x) {
  data.frame(
    y = 2^mean(log2(x)),
    ymin = 2^(mean(log2(x)) - 2 * log2(1.43)),
    ymax = 2^(mean(log2(x)) + 2 * log2(1.43))
  )
}

rprec_df = read_excel("paper_audit.xlsx", sheet = "Validation; Precision rubella") %>%
  mutate(ND50 = ND50 * RSV_A_CONV) %>%
  mutate(
    tick = case_when(
      `sample ID` == "R-02" ~ 1,
      `sample ID` == "R-01" ~ 2,
      `sample ID` == "R-08" ~ 3,
      `sample ID` == "R-12" ~ 4,
      `sample ID` == "R-14" ~ 5
    )
  )


rprec_df$logND50 = log2(rprec_df$ND50)

rprec_df = rprec_df %>%
  filter(plateID !="2023-Apr-25_plate5")



temp <- rprec_df %>%
  group_by(`sample ID`) %>%
  summarise(
    y = 2^mean(log2(ND50)),
    ymin = 2^(mean(log2(ND50)) - 2 * log2(1.43)),
    ymax = 2^(mean(log2(ND50)) + 2 * log2(1.43)),
    tick = mean(tick)
  )

r_prec = ggplot(rprec_df, aes(x = tick, y = ND50, group = tick)) +
  
  annotate(
    geom = "rect",
    xmin = temp$tick-0.3,
    xmax = temp$tick+0.3,
    ymax = temp$ymax,
    ymin = temp$ymin,
    alpha = 0.3,
    fill = 'green'
  ) +
  
  annotate(
    geom = "segment", 
    x = temp$tick - 0.3,
    xend =temp$tick + 0.3,
    y = temp$y,
    yend = temp$y,
    linetype = 'solid',
    size = 0.15
  ) + 
  
  geom_jitter(size = 1.5, alpha = 0.6, color = 'black', stroke = NA,
              width = 0.09,
              height = 0,
  ) + 

  scale_x_continuous(
    breaks = c(1, 2, 3, 4, 5),
    labels = c("R-02", "R-01", "R-08", "R-12", "R-14")
  ) + 

  scale_y_log10(
    breaks = c(1, 10, 100, 1000, 10000, 10000),
    labels = scales::trans_format("log10", scales::math_format(10^.x)),
    limits = c(1.55, 10^4)
  ) +
  
  
  theme_classic() +
  xlab('Panel member') + 
  ylab('Measured IU/mL') + 
  annotation_logticks(sides = 'l') + 
  
  theme(
    axis.text.x = element_text(size = 8, color = 'black'),
    axis.text.y = element_text(size = 8, color = 'black'),
    axis.title.x = element_text(size = 8), 
    axis.title.y = element_text(size = 8),
    plot.background = element_rect(fill = '#00000000'),
    panel.background = element_rect(fill = '#00000000'),
    panel.border = element_blank(),
    aspect.ratio = 1
  )

r_prec
```

## now combine panels into fig1 
```{r} 
valid_panel = plot_grid(fig_elisa, fig_linearity, l_prec, r_prec, labels = c('A', 'B', 'C', 'D'), axis = 'tblr', align = 'hv')
valid_panel
ggsave("paper_audit_figs/fig1.svg", valid_panel, dpi = 600, width = 5.5, height = 6)
```

```{r}
geo_mean <- function(x) {
  return (exp(mean(log(x))))
}
```

# Figure 2

```{r}
library(timetk)

df = read_excel("paper_audit.xlsx", sheet = "Results") %>%
  filter(!is.na(sample_type2))

df$ND50 = df$ND50 * RSV_A_CONV

timepoints = df %>%
  filter(sample_type == 'timepoint')

tp1 = timepoints %>%
  filter(CollDate < '2022/09/1')

tp2 = timepoints %>%
  filter_by_time(CollDate, '2022/07/01', '2023/02/01')

tp3 = timepoints %>%
  filter_by_time(CollDate, '2023/02/01', '2023/07/01')

tp4 = timepoints %>%
  filter(CollDate > '2023/07/01')


#### CREATE DATAFRAME FOR COMPARISON 

tp1$tp = 'timepoint 1'
tp2$tp = 'timepoint 2'
tp3$tp = 'timepoint 3'
tp4$tp = 'timepoint 4'

tp_comp = rbind(tp1, tp2, tp3, tp4)

my_groups = combn(unique(tp_comp$tp), 2)



lims <- as.POSIXct(strptime(c("2021-11-01 00:00", "2023-11-01 00:00"), 
                            format = "%Y-%m-%d %H:%M"))

fig2 = ggplot(timepoints, aes(x=CollDate, y=ND50)) + 
  geom_jitter(width = 19^5, size = 2, alpha = 0.54, stroke = 0.5) + 
  scale_y_log10(limits = c(15.5, 100000),
                breaks = c(10, 100, 1000, 10000, 100000),
                labels = scales::trans_format("log10", scales::math_format(10^.x)),) + 
  scale_x_datetime(date_breaks = ('1 month'), limits = lims) + 
  annotation_logticks(sides = 'l') + 
  
  #### geo_mean LINES 
  geom_segment(aes(x = as.POSIXct("2022-01-01"), xend = as.POSIXct("2022-04-01"), y = geo_mean(tp1$ND50), yend = geo_mean(tp1$ND50)), 
               linewidth = 0.5, 
               color = 'black') + 
  geom_segment(aes(x = as.POSIXct("2022-06-15"), xend = as.POSIXct("2022-09-15"), y = geo_mean(tp2$ND50), yend = geo_mean(tp2$ND50)), 
               linewidth = 0.5, 
               color = 'black') + 
  geom_segment(aes(x = as.POSIXct("2023-01-01"), xend = as.POSIXct("2023-03-20"), y = geo_mean(tp3$ND50), yend = geo_mean(tp3$ND50)), 
               linewidth = 0.5, 
               color = 'black') + 
  geom_segment(aes(x = as.POSIXct("2023-07-15"), xend = as.POSIXct("2023-10-15"), y = geo_mean(tp4$ND50), yend = geo_mean(tp4$ND50)), 
               linewidth = 0.5, 
               color = 'black') + 
  
  #### MEDIAN LINES 
  geom_segment(aes(x = as.POSIXct("2022-01-01"), xend = as.POSIXct("2022-04-01"), y = median(tp1$ND50), yend = median(tp1$ND50)), 
               linewidth = 0.5, 
               color = 'black',
               linetype = 'dashed') + 
  geom_segment(aes(x = as.POSIXct("2022-06-15"), xend = as.POSIXct("2022-09-15"), y = median(tp2$ND50), yend = median(tp2$ND50)), 
               linewidth = 0.5, 
               color = 'black',
               linetype = 'dashed') + 
  geom_segment(aes(x = as.POSIXct("2023-01-01"), xend = as.POSIXct("2023-03-20"), y = median(tp3$ND50), yend = median(tp3$ND50)), 
               linewidth = 0.5, 
               color = 'black',
               linetype = 'dashed') + 
  geom_segment(aes(x = as.POSIXct("2023-07-15"), xend = as.POSIXct("2023-10-15"), y = median(tp4$ND50), yend = median(tp4$ND50)), 
               linewidth = 0.5, 
               color = 'black',
               linetype = 'dashed') + 
  
  theme_classic() + 
  theme(axis.text.x = element_blank(),
        axis.text.y = element_text(size = 8, color = 'black'),
        axis.title.y = element_text(size = 9), 
        axis.title.x = element_blank(), 
        plot.background = element_rect(fill = '#00000000'),
        panel.background = element_rect(fill = '#00000000'),
        panel.border = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.position = 'none') + 
  ylab("Neutralizing titer (IU/mL)")

fig2


tests <- read_excel('paper_data/PCR_tests.xlsx')
colnames(tests) <- c('RepWeekDate', 'AntigenDetections')
tests$RepWeekDate <- as.Date(tests$RepWeekDate)


### MAKE FIGURE
fig1_overlay <- (ggplot(tests, aes(RepWeekDate, AntigenDetections, 
                        color=AntigenDetections)) + geom_line(linewidth=1.0, linetype="solid") + 
           scale_color_gradient(low="mediumseagreen", high="mediumseagreen")
) + 
  scale_y_continuous(position="left", limits=c(0, 800), 
                     breaks = c(0, 200, 400, 600, 800)) +
  
  coord_cartesian(xlim = as.Date(c('2021-11-01', '2023-11-01')), clip = 'on') + 
  
  theme(
    panel.background = element_rect(fill='transparent'),
    plot.background = element_rect(fill='transparent', color='transparent'),
    axis.text.x = element_text(size=7, angle=315, hjust=0, color = 'black'), 
    axis.text.y = element_text(size=8, color = 'black'),
    axis.title.y = element_text(size = 8, color = 'black'),
    axis.title.x = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    legend.background = element_blank(),
    legend.box.background = element_blank(),
    legend.text = element_blank(),
    panel.border = element_blank(),
    legend.position = "none",
    axis.line = element_line(color='black', linewidth=0.4),
    axis.ticks = element_line(color='black', linewidth=0.25) 
  ) + 
  guides(color=guide_legend(override.aes = list(color=NA))) + 
  
  scale_x_date(date_breaks=('1 month'), 
               # limits=as.Date(c('2021-11-01', '2023-11-01')),
               date_labels = "%b") + 
  ylab("# positive tests") 


fig1_overlay

fig2_total = ggarrange(fig2, fig1_overlay, 
                       nrow = 2, 
                       ncol = 1, 
                       heights = c(0.8, 0.2)
                       # align = "h"
                       )

fig2_total
ggsave("paper_audit_figs/fig2.svg", fig2_total, height = 5, width = 4.5, dpi = 600)
fig2_total

```

## Figure 3: clinical specimen groups:

```{r}
df = read_excel("paper_audit.xlsx", sheet = "Results") %>%
  filter(!is.na(sample_type2))


df$ND50 = df$ND50 * 0.37
df$logND50 = log2(df$ND50)

df$s2wrapped = str_wrap(df$sample_type2, width = 10, whitespace_only = TRUE)

my_groups = combn(unique(df$s2wrapped), 2)

compare = data.frame(
  df$s2wrapped,
  df$logND50
)

level_order = c("population",
                "future\nRSV-infected",
                "RSV+,\nimmunocompetent",
                "RSV+,\nimmunosuppressed",
                "Influenza+,\nRSV-"
                                 )

ic07 = df %>%
  filter(s2wrapped == "RSV+,\nimmunocompetent" & coll_period == '0-7' )

ic715 = df %>%
  filter(s2wrapped == "RSV+,\nimmunocompetent" & coll_period == '7-15' )

ic15 = df %>%
  filter(s2wrapped == "RSV+,\nimmunocompetent" & coll_period == '15+' )


fig3 = ggplot(df, aes(x=factor(s2wrapped,level = level_order), y=ND50 )) + 
  coord_cartesian(clip = 'off', ylim = c(19, 1*10^7)) + 
  theme_classic() + 
  geom_violin(alpha = 0.3, aes(fill = s2wrapped), color = "#00000000") +
  geom_beeswarm(size = 1, alpha = 0.8, cex = 1) +
  scale_y_log10(
                     breaks = c(10, 100, 1000, 10000, 100000),
                     labels = scales::trans_format("log10", scales::math_format(10^.x)),
                     limits = c(9, 10^6)
                ) +
  
  scale_x_discrete(labels = c("HSV serology\nremnant specimens\n(no PCR result)", 
                              "Future RSV PCR+\n(>10 days)", 
                              "RSV PCR+\nimmunocompetent\n(-7 - 98 days)", 
                              "RSV PCR+\nimmunosuppressed\n(-7 - 49 days)", 
                              "Influenza PCR+\nRSV PCR-\n(-7 - 39 days)")) + 
  
  annotation_logticks(sides = 'l') +
  
  geom_signif(
    comparisons = list(
    # my_groups[,1],
    my_groups[,8],
    my_groups[,2],
    # my_groups[,3],
    my_groups[,4],
    my_groups[,5],
    # my_groups[,6],
    # my_groups[,7],
    my_groups[,9]
    # my_groups[,10]
  ),
  test = "t.test",
  map_signif_level = TRUE,
  step_increase = 0.07,
  vjust = 0.8) +
  ylab("Neutralizing titer (IU/mL") + 

  theme(
    axis.text.x = element_text(size = 8, color = 'black'), 
    axis.text.y = element_text(size = 10, color = 'black'), 
    axis.title.y = element_text(size = 10, color = 'black'),
    axis.title.x = element_blank(),
    plot.background = element_rect(fill = '#00000000'),
    panel.background = element_rect(fill = '#00000000'),
    panel.border = element_blank(),
    legend.position = "none" 
  ) + 
  
## MEAN LINES
geom_segment(aes(x=0.6,
                 xend = 1.4,
                 y = geo_mean(subset(df, sample_type2 == 'population')$ND50),
                 yend = geo_mean(subset(df, sample_type2 == 'population')$ND50)),
             linewidth = 0.5,
             color = 'black') +

geom_segment(aes(x=4.6,
             xend = 5.4,
             y = geo_mean(subset(df, sample_type2 == 'Influenza+, RSV-')$ND50),
             yend = geo_mean(subset(df, sample_type2 == 'Influenza+, RSV-')$ND50)),
             linewidth = 0.5,
             color = 'black') +
geom_segment(aes(x=2.6,
                 xend = 3.4,
                 y = geo_mean(subset(df, sample_type2 == 'RSV+, immunocompetent')$ND50),
                 yend = geo_mean(subset(df, sample_type2 == 'RSV+, immunocompetent')$ND50)),
             linewidth = 0.5,
             color = 'black') +

geom_segment(aes(x=3.6,
                 xend = 4.4,
                 y = geo_mean(subset(df, sample_type2 == 'RSV+, immunosuppressed')$ND50),
                 yend = geo_mean(subset(df, sample_type2 == 'RSV+, immunosuppressed')$ND50)),
             linewidth = 0.5,
             color = 'black') +

geom_segment(aes(x=1.6,
                 xend = 2.4,
                 y = geo_mean(subset(df, sample_type2 == 'future RSV-infected')$ND50),
                 yend = geo_mean(subset(df, sample_type2 == 'future RSV-infected')$ND50)),
             linewidth = 0.5,
             color = 'black')

#           
fig3
ggsave("paper_audit_figs/fig3.svg", dpi = 600, width = 6, height = 5.2, unit = "in")
```
## Figure S1: neut titer and RSV infection time delta
```{r}
library(ggplot2)
library(readxl)
library(waterfalls)
library(tidyverse)
library(gridExtra)
library(ggpmisc)
library(scales)
library(dplyr)
library(gtable)
library(grid)
library(ggpattern)
library(gtable)
library(ggpubr)
library(svglite)
library(cowplot)
library(egg)

# df$ND50 = df$ND50 * 0.37

df = df %>% replace_na(list(performing_lab_locs = "HSV serology"))
df$hos_mask = df$performing_lab_locs

df_fil = df %>%
  filter(sample_type %in% c('recently_positive', 'flu', 'longitudinal', 'nan'))

df_fil = df_fil %>%
  filter(sample_type2 != 'HSV serology remnant specimens')

flu = df_fil %>%
  filter(sample_type2 == "Influenza+, RSV-"  )

rp_icompetent = df_fil %>%
  filter(sample_type2 == "RSV+, immunocompetent")

rp_isuppresssed = df_fil %>%
  filter(sample_type2 == "RSV+, immunosuppressed")

future_pos = df_fil %>%
  filter(sample_type2 == "future RSV-infected")

coords = coord_cartesian(clip = 'off', xlim = c(1, 100000)) 
age_overlay = geom_point(aes(x = 0.1, fill = `age_group`, , color = `age_group`, shape = hos_mask), size=1.3)
age_overlay_oof = geom_point(aes(x = 0.1, fill = `age_group`, , color = `age_group`), size=1.3)

age_overlay_altered = geom_point(aes(x = 0.1, shape= hos_mask), size=1.3) 
age_overlay_fp = geom_point(aes(x = 0.1, color = `age_group`), size=1.3) 
age_overlay_hos = geom_point(aes(x = 0.1, shape = hos_mask), size = 1.3)

hos_scale = scale_shape_manual(
  name = '',
  values = c('Hospital A'=24, 'Hospital B'=22, 'Hospital C'=23, 'HSV serology' = 25 )
) 

age_scale_fill = scale_fill_manual(
  name = '',
  values = c('65+'='darkslategray', 
             '18-65'='darkolivegreen4', 
             '18 and below'='darkseagreen1', 
             'Unknown'='dimgrey'),
  labels = c('18 years and below', '18-65 years', '65+ years', 'unknown') 
)

age_scale_color = scale_color_manual(
  name = '',
  values = c('65+'='darkslategray', 
             '18-65'='darkolivegreen4', 
             '18 and below'='darkseagreen1', 
             'Unknown'='dimgrey'),
  labels = c('18 years and below', '18-65 years', '65+ years', 'unknown') 
)

level_order = c('future RSV-infected', 'RSV+, immunocompetent', 'RSV+, immunosuppressed', 'Influenza+, RSV-')
### COLOR LEGEND
color_plot = ggplot(df_fil, aes(x=ND50, y = reorder(dummy, ND50), fill = factor(sample_type2, level = level_order))) +
  geom_col() + 
  theme_classic() + 
  scale_y_discrete(position = 'left') +
  scale_x_log10(
    breaks = scales::trans_breaks("log10", function(x) 10^x),
    labels = scales::trans_format("log10", scales::math_format(10^.x)),
    # limits = c(1, 100000)
  ) +
  scale_fill_manual(name = "",
    values = c(
      "future RSV-infected" = "mediumpurple",
      "RSV+, immunocompetent"="brown", 
      "RSV+, immunosuppressed"="coral",
      "Influenza+, RSV-" = "blue"
      
    ),
    
    labels = c(
                        "Future RSV PCR+\n(>10 days)", 
                        "RSV PCR+\nimmunocompetent\n(-7 - 98 days)", 
                        "RSV PCR+\nimmunosuppressed\n(-7 - 49 days)", 
                        "Influenza PCR+\nRSV PCR-\n(-7 - 39 days)")
  ) + 
  annotation_logticks(sides='b') +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_text(size=6),
        axis.text.x = element_text(size=8, color = 'black'),
        axis.title.x = element_blank(),
        legend.position = "right",
        # legend.direction = "horizontal",
        legend.title = element_blank(),
        legend.text = element_text(size = 6),

  ) + 
 guides(fill = guide_legend(ncol = 4, nrow = 1))


color_plot 

color_legend = cowplot::get_legend(color_plot)

color_legend

age_plot = ggplot(df_fil, aes(x=ND50, y = reorder(dummy, ND50))) +
  coords + 
  age_overlay_oof +
  age_scale_fill + 
  age_scale_color+ 
  geom_col() + 
  theme_classic() + 
  scale_y_discrete(position = 'left') +
  scale_x_log10(
    breaks = scales::trans_breaks("log10", function(x) 10^x),
    labels = scales::trans_format("log10", scales::math_format(10^.x))
  ) +

  theme(axis.title.y = element_blank(),
        axis.text.y = element_text(size=6),
        axis.text.x = element_text(size=8, color = 'black'),
        axis.title.x = element_blank(),
        legend.position = "right",
        legend.title = element_blank(),
        legend.text = element_text(size = 6, color = "black"),

  ) + 
  guides( fill = guide_legend(nrow =1, ncol =4 ), label.hjust = -0.4)

age_plot

age_legend = cowplot::get_legend(age_plot)


##### GET LEGEND FOR HOSPITALS
hos_plot = ggplot(df_fil, aes(x=ND50, y = reorder(dummy, ND50))) +
  coords + 
  age_overlay_hos +
  hos_scale + 
  geom_col() + 
  theme_classic() + 
  scale_y_discrete(position = 'left') +
  scale_x_log10(
    breaks = scales::trans_breaks("log10", function(x) 10^x),
    labels = scales::trans_format("log10", scales::math_format(10^.x))
  ) +
  
  theme(axis.title.y = element_blank(),
        axis.text.y = element_text(size=6),
        axis.text.x = element_text(size=8, color = 'black'),
        axis.title.x = element_blank(),
        legend.position = "right",
        legend.title = element_blank(),
        legend.text = element_text(size = 6),
        
  ) + 
  guides( shape = guide_legend(nrow =1, ncol =4 ), label.hjust = -0.4)


hos_plot

hos_legend = cowplot::get_legend(hos_plot)

hos_legend

### CUBE ROOT TRANSFORM 
cube_root <- function(x) sign(x) * abs(x) ^ (1/3)
cube <- function(x) x ^ 3
library(scales)
trans_cube = trans_new(name = "cube root", 
                       transform = cube_root, 
                       inverse = cube)

##### FLU timedelta
flu_dt = ggplot(flu, aes(x=collected_days_since_pos, y = `sample ID`)) +
  geom_col(fill = 'blue') +
  theme_classic() + 
  scale_y_discrete(position='right') + 
  scale_x_continuous(
    trans = trans_cube,
    breaks = c(-500, -120, -30, -5, 0, 5, 30, 120),
    limits = c(-500, 120)
  ) + 
  theme_classic() +
  theme(axis.text.y = element_blank(),
        axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_text(size=8, angle = 0, color = 'black'),
        legend.position="right",
        legend.text = element_text(size=6), legend.spacing = unit(0.1, 'points'),
        plot.background = element_rect(fill = '#00000000'),
        panel.background = element_rect(fill = '#00000000'),
        panel.border = element_blank(),
        

  )

flu_dt

### recently positive immunocompetent
rp_icompetent_dt = ggplot(rp_icompetent, aes(x=collected_days_since_pos, y = reorder(dummy, ND50))) +
  geom_col(fill = 'brown') +
  theme_classic() + 
  scale_y_discrete(position='right') + 
  scale_x_continuous(
    trans = trans_cube,
    limits = c(-500, 120),
    breaks = c(-500, -120, -30, -5, 0, 5, 30, 120)
  ) + 
  theme_classic() +
  theme(axis.text.y = element_blank(),
        axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_text(size=8, angle = 0, color = 'black'),
        legend.position="right",
        legend.text = element_text(size=6), legend.spacing = unit(0.1, 'points'),
        plot.background = element_rect(fill = '#00000000'),
        panel.background = element_rect(fill = '#00000000'),
        panel.border = element_blank(),

  )

rp_icompetent_dt

### recently positive immmunosuppressed timedelta
rp_isuppressed_dt = ggplot(rp_isuppresssed, aes(x=collected_days_since_pos, y = reorder(dummy, ND50))) +
  geom_col(fill = 'coral') +
  theme_classic() + 
  scale_y_discrete(position='right') + 
  scale_x_continuous(
    trans = trans_cube,
    limits = c(-500, 120),
    breaks = c(-500, -120, -30, -5, 0, 5, 30, 120)
  ) + 
  theme_classic() +
  theme(axis.text.y = element_blank(),
        axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_text(size=8, angle = 0, color = 'black'),
        legend.position="right",
        legend.text = element_text(size=6), legend.spacing = unit(0.1, 'points'),
        plot.background = element_rect(fill = '#00000000'),
        panel.background = element_rect(fill = '#00000000'),
        panel.border = element_blank(),

  )

rp_isuppressed_dt

### future-infected timedelta 

future_pos_dt = ggplot(future_pos, aes(x=collected_days_since_pos, y = reorder(dummy, ND50))) +
  geom_col(fill = 'mediumpurple') +
  theme_classic() + 
  scale_y_discrete(position='right') + 
  scale_x_continuous(
    trans = trans_cube,
    breaks = c(-500, -120, -30, -5, 0, 5, 30, 120),
    limits = c(-500, 120)
  ) + 
  theme_classic() +
  theme(axis.text.y = element_blank(),
        axis.title.y = element_blank(),
        axis.title.x = element_text(size = 10),
        axis.text.x = element_text(size=8, angle = 0, color = 'black'),
        legend.position="right",
        legend.text = element_text(size=6), legend.spacing = unit(0.1, 'points'),
        plot.background = element_rect(fill = '#00000000'),
        panel.background = element_rect(fill = '#00000000'),
        panel.border = element_blank(),

  ) + 
  xlab("Serum draw date - PCR date")

future_pos_dt

### flu titer 

flu_titer = ggplot(flu, aes(x=ND50, y = reorder(dummy, ND50))) +
  coords + 
  age_overlay + 
  age_scale_color +
  age_scale_fill + 
  hos_scale + 
  geom_col(fill='blue') + 
  theme_classic() + 
  scale_y_discrete(position = 'left') +
  scale_x_log10(
    breaks = scales::trans_breaks("log10", function(x) 10^x),
    labels = scales::trans_format("log10", scales::math_format(10^.x))
  ) +
  annotation_logticks(sides='b') +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_text(size=6),
        axis.text.x = element_text(size=8, color = 'black'),
        axis.title.x = element_blank(),
        plot.background = element_rect(fill = '#00000000'),
        panel.background = element_rect(fill = '#00000000'),
        panel.border = element_blank(),
        legend.position = "none",

  )    

flu_titer

rp_icompetent_titer = ggplot(rp_icompetent, aes(x=ND50, y = reorder(dummy, ND50))) +
  coords + 
  geom_col(fill = 'brown') + 
  age_overlay + 
  age_scale_color +
  age_scale_fill + 
  hos_scale + 
  theme_classic() + 
  scale_y_discrete(position = 'left') +
  scale_x_log10(
    breaks = scales::trans_breaks("log10", function(x) 10^x),
    labels = scales::trans_format("log10", scales::math_format(10^.x))
  ) +
  annotation_logticks(sides='b') +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_text(size=6),
        axis.text.x = element_text(size=8, color = 'black'),
        axis.title.x = element_blank(),
        plot.background = element_rect(fill = '#00000000'),
        panel.background = element_rect(fill = '#00000000'),
        panel.border = element_blank(),
        legend.position = "none",

  )  

rp_icompetent_titer
  
rp_isuppresssed_titer = ggplot(rp_isuppresssed, aes(x=ND50, y = reorder(dummy, ND50))) +
  coords + 
  age_overlay + 
  age_scale_fill +
  age_scale_color + 
  hos_scale + 
  geom_col(fill = 'coral') + 
  theme_classic() + 
  scale_y_discrete(position = 'left') +
  scale_x_log10(
    breaks = scales::trans_breaks("log10", function(x) 10^x),
    labels = scales::trans_format("log10", scales::math_format(10^.x)),
    # limits = c(1, 100000)
  ) +
  annotation_logticks(sides='b') +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_text(size=6),
        axis.text.x = element_text(size=8, color = 'black'),
        axis.title.x = element_blank(),
        plot.background = element_rect(fill = '#00000000'),
        panel.background = element_rect(fill = '#00000000'),
        panel.border = element_blank(),
        legend.position = "none",

  )  

rp_isuppresssed_titer

future_pos_titer = ggplot(future_pos, aes(x=ND50, y = reorder(dummy, ND50))) +
  coords + 
  age_overlay + 
  age_scale_fill +
  age_scale_color + 
  hos_scale + 
  geom_col(fill = 'mediumpurple') + 
  theme_classic() + 
  scale_y_discrete(position = 'left') +
  scale_x_log10(
    breaks = scales::trans_breaks("log10", function(x) 10^x),
    labels = scales::trans_format("log10", scales::math_format(10^.x)),
    # limits = c(1, 100000)
  ) +
  annotation_logticks(sides='b') +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_text(size=6),
        axis.text.x = element_text(size=8, color = 'black'),
        axis.title.x = element_text(size = 10),
        plot.background = element_rect(fill = '#00000000'),
        panel.background = element_rect(fill = '#00000000'),
        panel.border = element_blank(),
        legend.position = "none"

  ) +
  xlab("Neutralizing titer (IU/ml)")

future_pos_titer


age_legend = ggdraw(age_legend) +
  theme(
    legend.background = element_rect(fill = "transparent", color = NA),
    legend.box.background = element_rect(fill = "transparent", color = NA),
    legend.key = element_rect(fill = "transparent", color = NA),
    panel.background = element_rect(fill = "transparent", color = NA),
    plot.background = element_rect(fill = "transparent", color = NA)
  ) 
        

hos_legend = ggdraw(hos_legend) + 
  theme(
    legend.background = element_rect(fill = "transparent", color = NA),
    legend.box.background = element_rect(fill = "transparent", color = NA),
    legend.key = element_rect(fill = "transparent", color = NA),
    panel.background = element_rect(fill = "transparent", color = NA),
    plot.background = element_rect(fill = "transparent", color = NA)
  )

color_legend = ggdraw(color_legend) + 
  theme(
    legend.background = element_rect(fill = "transparent", color = NA),
    legend.box.background = element_rect(fill = "transparent", color = NA),
    legend.key = element_rect(fill = "transparent", color = NA),
    panel.background = element_rect(fill = "transparent", color = NA),
    plot.background = element_rect(fill = "transparent", color = NA)
  )

alt_grid = ggarrange(

                     flu_dt,
                     flu_titer,
                     rp_icompetent_dt,
                     rp_icompetent_titer,
                     rp_isuppressed_dt,
                     rp_isuppresssed_titer,
                     future_pos_dt,
                     future_pos_titer,
                     nrow = 4,
                     ncol = 2,
                     heights = c(0.7, 1.3, 0.7, 1.3)
                     # align = 'hv'
)


alt_grid

legends = egg::ggarrange(color_legend, 
                    age_legend,
                    hos_legend,
                    nrow = 3, ncol = 1,
                    
                    heights = c(1, 1, 1),
                    padding = unit(2, "mm"))


ggsave("figs2_legends.svg", legends, dpi = 1000)

full_grid = cowplot::plot_grid(alt_grid, 
                      legends,
                     nrow = 2,
                     ncol = 1,
                     rel_heights = c(1.80, 0.2))
full_grid

ggsave(filename = 'paper_audit_figs/figS1.svg',
       plot = full_grid,
       dpi = 1000,
       width = 5.6,
       height = 8.2,
       bg = "transparent"
       )


```

```{r}

geo_mean <- function(x) {
  return (exp(mean(log(x))))
}

geomSD <- function(x) {
  sd_log_x <- exp(sd(log(x)))
  
  return(sd_log_x)
}

calculate_GCV <- function(sd_log_data) {
  # Calculate GCV from standard deviation of log-transformed data, used for anovaVCA output
  gcv <- sqrt(exp(sd_log_data^2) - 1) * 100
  
  # Return the result
  return(gcv)
}

```

```{r}
calculate_variance_components <- function(anova_tab, measurement_df, measurement_col, sample_id_col) {
  lapply(names(anova_tab), function(sample_id) {
    print(sample_id)
    vca_obj <- anova_tab[[sample_id]]
    print(vca_obj)
    tab = as.data.frame(vca_obj[["aov.tab"]])
    
    br_v = (tab[["MS"]][2] - tab[["MS"]][3]) / 2
    wr_v = tab[["MS"]][3]
    tot_v = br_v + wr_v
    
    br_sd = sqrt(br_v)
    wr_sd = sqrt(wr_v)
    tot_sd = sqrt(tot_v)
    
    br_cv = calculate_GCV(br_sd)
    wr_cv = calculate_GCV(wr_sd)
    tot_cv = calculate_GCV(tot_sd)
    
    gmeans = measurement_df %>%
      group_by(!!sym(sample_id_col)) %>%
      reframe(!!sym(sample_id_col) := !!sym(sample_id_col), 
              geo_mean = geo_mean(!!sym(measurement_col))) %>%
      distinct()
    
    result = data.frame(
      measurement = measurement_col,
      panel_member = sample_id,
      intraassay_GSD = wr_sd,
      intraassay_GCV = wr_cv,
      interassay_GSD = br_sd,
      interassay_GCV = br_cv,
      total_GSD = tot_sd,
      total_GCV = tot_cv
    )
    
    print(result)
    
    left_join(result, gmeans, by = sample_id_col)
  }) %>% bind_rows()
}
```

# table S1
```{r}

a_conv_df = read_excel("paper_audit.xlsx", sheet = "RSV A conversion factor data") %>%
  mutate(virus = "RSV A2") %>%
  pivot_longer(
    cols = c(ND50, ND80), 
    names_to = "measurement",
    values_to = "titer") %>%
  mutate(log_titer = log(titer)) %>%
  mutate(day = case_when(
    str_detect(plateID, "May-15") ~ 1,
    str_detect(plateID, "May-16") ~ 2, 
    str_detect(plateID, "May-22") ~ 3,
    str_detect(plateID, "May-23") ~ 4
  )) %>%
  group_by(measurement) %>%
  mutate(panel_member = `sample ID`, virus = virus, measurement = measurement, geomean = geo_mean(titer), conversion_factor = 2000 / geomean) %>% 
    distinct()

a_conv_50 = anovaVCA(log_titer ~ day,
                     Data = as.data.frame(a_conv_df %>% filter(measurement == "ND50")),
                     by = "sample ID",
                     NegVC = TRUE)

b_conv_df = read_excel("paper_audit.xlsx", sheet = "Validation; RSV B") %>%
  filter(`sample ID` == "NIBSC 16.284") %>%
  mutate(virus = "RSV B WV/14617") %>%
  pivot_longer(
    cols = c(ND50, ND80), 
    names_to = "measurement",
    values_to = "titer") %>%
  mutate(log_titer = log(titer)) %>%
  mutate(day = case_when(
    str_detect(plateID, "Nov-27") ~ 1,
    str_detect(plateID, "Dec_4") ~ 2, 
    str_detect(plateID, "Dec-5") ~ 3,
  )) %>%
  group_by(measurement) %>%
  summarise(virus = virus, measurement = measurement, geo_mean = geo_mean(titer), conversion_factor = 2000 / geo_mean) %>%
    distinct()

convf_df = bind_rows(a_conv_df, b_conv_df)
```

JR - sensitivity analysis

```{r}
library(tidyverse)
library(scales)

# rm(list=ls())

runs <- c("RSV/expansions and titer/working stock titers/2023-Apr-04 (neut & RSV A2 working stock titrat)/neut",
          "RSV/2023-Apr-10 (neut)", 
          "RSV/2023-Apr-17 (nuet)", 
          "RSV/2023-Apr-24 (Neut)",
          "RSV/2023-Apr-25 (neut)", 
          "RSV/2023-May-01 (neut)", 
          "RSV/2023-May-03 (neut)",
          "RSV/2023-May-09 (neut)", 
          "RSV/2023-May-15 (neut-cursed)", 
          "RSV/2023-May-16 (neut)",
          "RSV/2023-May-22 (neut)", 
          "Cascadia/runs/2023-May-23 EP (neut)/dr4pl analysis/")

dat <- tibble(path = runs) %>%
  mutate(
    file = map(path, ~list.files(.x, pattern = "\\d\\d\\d\\d.+\\.csv"))
  ) %>%
  unnest(file) %>%
  filter(
    !str_detect(file, "Results")
  ) %>%
  mutate(
    data = map(file.path(path, file), read_csv)
  ) %>%
  unnest(data)
    

dat <- dat %>%
  mutate(
    run_date = str_extract(plateID, "\\d\\d\\d\\d-...-\\d\\d"),
    .after = plateID
  )

#filter out excluded runs

dat_filter <- dat %>%
  group_by(plateID) %>%
  mutate(VOC_avg = mean(foci_num[type=="VOC"], na.rm = TRUE)) %>%
  mutate(neg_avg = mean(foci_num[type=="negative"], na.rm = TRUE)) %>%
  group_by(plateID, sample_num) %>%
  mutate(foci_red = 1-(foci_num/VOC_avg)) %>%
  
  #(2 - 6) 2023-Apr-04 plate 1,2, run used row format
  filter(!(run_date == "2023-Apr-04" & str_detect(plateID, "plate1|plate2"))) %>%
  
  #(1 - 6) 2023-Apr-10 plate 2, run used row format
  filter(!(run_date == "2023-Apr-10" & str_detect(plateID, "plate2"))) %>%
  
  #excluded - stock comparisons and pipet var.
  #(1 - 24) 2023-May-16 plate 1 old RSV A2
  #(1 - 24) 2023-May-16 plate 2 new RSV A2 working stock test dil #1
  #(1 - 24) 2023-May-16 plate 3 new RSV A2 working stock test dil #2
  #(1 - 24) 2023-May-16 plate 4 old RSV A2 working stock pipet variation
  filter(!(run_date == "2023-May-16" & str_detect(plateID, "plate1|plate2|plate3|plate4"))) %>%
  
  #(1 - 24) 2023-May-22 plate 1 high/low/neg control check - low foci counts
  #(1 - 24) 2023-May-22 plate 2 pipet variation - low foci counts
  filter(!(run_date == "2023-May-22" & str_detect(plateID, "plate1|plate2"))) %>%
  
  #(1 - 12) 2023-May-23 plate 5 cascadia run
  #(1 - 12) 2023-May-23 plate 6 cascadia run
  filter(!(run_date == "2023-May-23" & str_detect(plateID, "plate5|plate6")))

#temp_voc <- dat_filter %>% filter(type == "VOC") %>% group_by(plateID) %>% summarise(count = n())
#temp_neg <- dat_filter %>% filter(type == "negative") %>% group_by(plateID) %>% summarise(count = n())

###############included runs:

#12 VOC runs
#(1 - 12) 2023-Apr-10 plate 1 (NR-4023) BEI controls
#(1 - 12) 2023-Apr-17 plate 1 BEI controls
#(6 - 12) 2023-Apr-24 plate 1,2,3,4,5,6 BEI controls and Rubella specimens
#(5 - 12) 2023-Apr-25 plate 1 (NR-4203),2,3,4,5 BEI controls and lin specimens only
#(5 - 12) 2023-May-01 plate 1 (NR-4023),2,3,4,5 BEI controls, lin, and rubella specimens only
#(1 - 12) 2023-May-01 plate 6 (NR-4023) 26 hpi infection test, with NR-4023
#(5 - 12) 2023-May-03 plate 1 (NR-4023 x 2),2,3,4,5 BEI controls, lin and rubella specimens only
#(1 - 12) 2023-May-03 plate 6 (NR-4023 x 2) 26 hpi infection test, with NR-4023
#(1 - 12) 2023-May-09 plate 1 BEI controls
#(1 - 12) 2023-May-15, plate 1, run was labeled cursed, NIBSC stocks
#(1 - 12) 2023-May-16 plate 5 NIBSC stocks only, old RSV A2
#(2 - 12) 2023-May-22 plate 3,4 NIBSC and BEI replicates
#(2 - 12) 2023-May-23 plate 3, 4 NIBSC BEI controls

#24 VOC runs
#(1 - 24) 2023-May-23 plate1 high/low/neg control only
#(1 - 24) 2023-May-23 plate2 high/low/neg control check - duplicate for shah

#48 VOC runs
#(1 - 48) 2023-May-09 plate 2 negative only & VOC


dat_VOC_stat <- dat_filter %>%
  ungroup %>%
  filter(type == "VOC") %>%
  summarise(
    avg_VOC = mean(VOC_avg),
    sd_VOC = sd(VOC_avg),
    min_VOC = min(VOC_avg),
    max_VOC = max(VOC_avg)
  )

dat_sum <- dat_filter %>%
  filter(type == "VOC" | type == "negative" | type == "NR-4023 anti-RSV pooled serum – low") %>%
  group_by(type, fold_dil) %>%
  summarise(
    avg = mean(foci_red),
    sd = sd(foci_red),
    count = n()
  )

LOB_lower <- dat_sum %>%
  filter(type == "VOC") %>%
  mutate(LOB = avg + 1.645*sd)

LOB_upper <- dat_sum %>%
  filter(type == "negative") %>%
  mutate(LOB = avg - 1.645*sd)

LOD_lower <- dat_sum %>%
  filter(fold_dil == 24300) %>%
  mutate(LOD_lower = LOB_lower$LOB + 1.645*sd)

LOD_upper <- dat_sum %>%
  filter(fold_dil == 300) %>%
  mutate(LOD_upper = LOB_upper$LOB - 1.645*sd)



tab_S3 <- tribble(~`Assay parameter`, ~Lower, ~Upper,
                  "LoB", LOB_lower$LOB, LOB_upper$LOB,
                  "LoD", LOD_lower$LOD_lower, LOD_upper$LOD_upper) %>%
  mutate(
    Lower = percent(Lower, accuracy = 0.1),
    Upper = percent(Upper, accuracy = 0.1)
  ) %>%
  flextable()

tab_S3

tab_S4 <- tribble(~control, ~`Mean (% inhibition)`, ~`Standard deviation`, ~n,
                  "Virus only control (VOC)", LOB_lower$avg, LOB_lower$sd, LOB_lower$count,
                  "Media only control", LOB_upper$avg, LOB_upper$sd, LOB_upper$count,
                  "NR-4023 24,300-fold diluted", LOD_lower$avg, LOD_lower$sd, LOD_lower$count,
                  "NR-4023 300-fold diluted", LOD_upper$avg, LOD_upper$sd, LOD_upper$count) %>%
  mutate(
    `Mean (% inhibition)` = percent(`Mean (% inhibition)`, accuracy = 0.1),
    `Standard deviation`= percent(`Standard deviation`, accuracy = 0.1)
  ) %>%
  flextable()

tab_S4

rm(LOB_lower, LOB_upper, LOD_lower, LOD_upper, runs, dat, dat_filter,
   dat_VOC_stat, dat_sum)


```

# table S5
```{r}
tab_s5_df = read_excel("paper_audit.xlsx", sheet = "Table S5")

tab_s5_df = tab_s5_df %>%
  mutate(
    dil_factor = case_when(
      `sample ID` == "lin #2" ~ 8,
      `sample ID` == "lin #3" ~ 32,
      `sample ID` == "lin #4" ~ 128,
      `sample ID` == "lin #5" ~ 512), 
    pre_dilution = paste0("1:", dil_factor),
    IU = 0.37 * ND50
  ) %>% 
  group_by(`sample ID`) %>%
  summarise(
    n = n(),
    measured_IU = mean(IU), 
    IU_predil = measured_IU * dil_factor,
    pre_dilution = pre_dilution,
    ratio_expected = round(IU_predil / (29663 * 0.37), 2),
    IU_predil = round(IU_predil, 0),
    measured_IU = round(measured_IU, 0)
  ) %>% ungroup() %>% distinct()

tab_s5 = tab_s5_df %>%
  select(pre_dilution, n, measured_IU, IU_predil, ratio_expected) %>%
  flextable() %>%
  align(part = "all", align = "center")
  

tab_s5
```
# table 1
```{r}
library(flextable)

tab_1_df = read_excel("paper_audit.xlsx", sheet = "Table 1")
tab_1_sum = tab_1_df %>%
  group_by(`sample ID`) %>%
  summarise(count = n(), geo_mean = geo_mean(ND50), mean_IU = geo_mean * 0.37, mean_IU_rounded = round(mean_IU, 0)) %>%
  mutate(reference = round(case_when(
    `sample ID` == "NR-21973" ~ 6830,
    `sample ID` == "NR-4020" ~ 1236,
    `sample ID` == "NR-4021" ~ 4404,
    `sample ID` == "NR-4022" ~ 706,
    `sample ID` == "NR-4023" ~ 625,
  ), 1)) %>%
  mutate(ratio = round(mean_IU / reference, 1))

tab_1 = tab_1_sum %>%
  filter(`sample ID` != "NIBSC 16/284") %>%
  select(c(`sample ID`, count, mean_IU_rounded, reference, ratio)) %>%
  rename(c(
    sample = "sample ID", 
    N = "count",
    `IU/mL RSV FRNT` = "mean_IU_rounded", 
    `IU/mL reference` = "reference", 
     `Ratio of IU/mL\nRSV FRNT/ reference` = "ratio"
    )) %>%
  flextable() %>%
  align(part = "all", align = "center") %>%
  autofit()

tab_1 %>% save_as_pptx(path = "paper_audit_figs/tab_1.pptx")
tab_1 
```
# table 2
```{r} 
tab_2_df = read_excel("paper_audit.xlsx", sheet = "Validation; RSV B") %>%
  filter(`sample ID` %in% c("NR-21973", "NR-4022")) %>%
  group_by(`sample ID`) %>%
  summarise(count = n(), mean = round(mean(ND50 * 0.91), 0), geo_mean = geo_mean(ND50), mean_IU = geo_mean * 0.91, mean_IU_rounded = round(mean_IU, 0)) %>%
  mutate(reference = round(case_when(
    `sample ID` == "NR-21973" ~ 12059,
    `sample ID` == "NR-4022" ~ 1005,
  ), 1)) %>%
  mutate(ratio = round(mean_IU / reference, 1))

tab_2 = tab_2_df %>%
  select(c(`sample ID`, count, mean_IU_rounded, reference, ratio)) %>%
  rename(c(
    sample = "sample ID", 
    N = "count",
    `IU/mL RSV FRNT` = "mean_IU_rounded",
    `IU/mL reference` = "reference",
    `Ratio of IU/mL\nRSV FRNT/ reference` = "ratio"
  )) %>%
  flextable() %>%
  align(part = "all", align = "center") %>%
  autofit()

tab_2
```
# table 3: contrived precision

NOTE ON EXPECTED MEASUREMENT CALCULATIONS:
expected measurements are calculated by converting the average ND50/ND80 of each panel member to IU/mL. Average values (obtained from the 2023-Apr-10 run) can be found in `RSV assay validation v3.docx`; a copy of which should be in `RSV/Qualification/Qualification report/`. 

Expected measurements can be found on page 8 of the aforementioned document, and are shown below: 

specimen	                    ND50	  ND80    Dilution (cumulative)
-------------------------------------------------------------------
NR-21973 anti-RSV ref. serum	29,963	7,652	  neat
specimen #1	                  14,982	3,826	  2 (2)
specimen #2	                  3,745	  957	    4 (8)
specimen #3	                  936	    239	    4 (32)
specimen #4	                  234	    60	    4 (128)
specimen #5	                  59	    15	    4 (512)
specimen #6	                  15	    4	      4 (2048)


```{r}
tab_3_df = read_excel("paper_audit.xlsx", sheet = "Validation; Precision contrived") %>%
  filter(`sample ID` != "negative control") %>%
  group_by(`sample ID`) %>%
  mutate(panel_member = `sample ID`, sid = as.integer(str_split_i(panel_member, "lin #", 2))) %>%
  ungroup() %>%
  group_by(panel_member) %>%
  mutate(
    n = n(), 
    ## ND50
    IU_50 = ND50 * 0.37,
    logIU_50 = log(IU_50),
    # `ND50__geomean` = geo_mean(IU_50),
    ## ND80
    IU_80 = ND80 * 1.32, 
    logIU_80 = log(IU_80),
    # `ND80__geomean` = geo_mean(IU_80),
    ) %>%
  ungroup() %>%
  mutate(day = case_when(
    str_detect(run_date, "04-25") ~ 1,
    str_detect(run_date, "05-01") ~ 2,
    str_detect(run_date, "05-03") ~ 3,
  )) %>%
  group_by(day) %>%
  mutate(plate_num = dense_rank(plateID)) %>%
  ungroup()

anova_tab_50 = anovaVCA(logIU_50 ~ day,
                      Data = as.data.frame(tab_3_df),
                      by = "sample ID",
                      NegVC = TRUE)

tab_3_anova_50 <- lapply(names(anova_tab_50), function(sample_id) {
  vca_obj <- anova_tab_50[[sample_id]]
  tab = as.data.frame(vca_obj[["aov.tab"]])
  
  br_v = (tab[["MS"]][2] - tab[["MS"]][3]) / 2
  wr_v = tab[["MS"]][3]
  tot_v = br_v + wr_v
  
  br_sd = sqrt(br_v)
  wr_sd = sqrt(wr_v)
  tot_sd = sqrt(tot_v)
  
  br_cv = calculate_GCV(br_sd)
  wr_cv = calculate_GCV(wr_sd)
  tot_cv = calculate_GCV(tot_sd)
  
  gmeans = tab_3_df %>%
    group_by(panel_member) %>%
    reframe(panel_member = panel_member, mean = mean(IU_50), geo_mean = geo_mean(IU_50)) %>%
    distinct()
  
  left_join(data.frame(
    measurement = "ND50",
    panel_member = str_split_i(sample_id, "sample ID.", 2),
    
    intraassay_GSD = wr_sd,
    intraassay_GCV = wr_cv,
    
    interassay_GSD = br_sd,
    interassay_GCV = br_cv,
    
    total_GSD = tot_sd,
    total_GCV = tot_cv
  ) %>% mutate(
    expected_measurement = case_when(
      # panel_member == "lin #2" ~ 1386,
      # panel_member == "lin #3" ~ 346,
      # panel_member == "lin #4" ~ 87,
      # panel_member == "lin #5" ~ 22,
      panel_member == "lin #2" ~ round(29963 / 8 * 0.37, 0),
      panel_member == "lin #3" ~ round(29963 / 32 * 0.37, 0),
      panel_member == "lin #4" ~ round(29963 / 128 * 0.37, 0),
      panel_member == "lin #5" ~ round(29963 / 512 * 0.37, 0)
    )), gmeans, by = "panel_member")
})

anova_tab_80 = anovaVCA(logIU_80 ~ day,
                      Data = as.data.frame(tab_3_df),
                      by = "sample ID",
                      NegVC = TRUE)

tab_3_anova_80 <- lapply(names(anova_tab_80), function(sample_id) {
  vca_obj <- anova_tab_80[[sample_id]]
  tab = as.data.frame(vca_obj[["aov.tab"]])
  
  br_v = (tab[["MS"]][2] - tab[["MS"]][3]) / 2
  wr_v = tab[["MS"]][3]
  tot_v = br_v + wr_v
  
  br_sd = sqrt(br_v)
  wr_sd = sqrt(wr_v)
  tot_sd = sqrt(tot_v)
  
  br_cv = calculate_GCV(br_sd)
  wr_cv = calculate_GCV(wr_sd)
  tot_cv = calculate_GCV(tot_sd)
  
  gmeans = tab_3_df %>%
    group_by(panel_member) %>%
    reframe(panel_member = panel_member, mean = mean(IU_80), geo_mean = geo_mean(IU_80)) %>%
    distinct()
  
  left_join(data.frame(
    measurement = "ND80",
    panel_member = str_split_i(sample_id, "sample ID.", 2),
    
    intraassay_GSD = wr_sd,
    intraassay_GCV = wr_cv,
    
    interassay_GSD = br_sd,
    interassay_GCV = br_cv,
    
    total_GSD = tot_sd,
    total_GCV = tot_cv
  ) %>% 
    mutate(
    expected_measurement = case_when(
      # panel_member == "lin #2" ~ 1263,
      # panel_member == "lin #3" ~ 315,
      # panel_member == "lin #4" ~ 79,
      # panel_member == "lin #5" ~ 2,
      panel_member == "lin #2" ~ round(7652 / 8 * 1.32, 0),
      panel_member == "lin #3" ~ round(7652 / 32 * 1.32, 0),
      panel_member == "lin #4" ~ round(7652 / 128 * 1.32, 0),
      panel_member == "lin #5" ~ round(7652 / 512 * 1.32, 0)
    )),
  gmeans, by = "panel_member")
})

anova_df = bind_rows(tab_3_anova_50, tab_3_anova_80)


tab_3_df = left_join(tab_3_df, anova_df, by = "panel_member")

tab_3 = tab_3_df %>%
  select(
    measurement, 
    panel_member, 
    expected_measurement,
    n, 
    # geo_mean,
    mean,
    matches(".*_GSD|.*_GCV"), 
    matches(".*_GSD|.*_GCV")) %>%
  distinct() %>%
  rename_with(
    .cols = matches("intraassay"),
    .fn = ~str_replace(., "intraassay", "Intra-assay component")
  ) %>%
  rename_with(
    .cols = matches("interassay"),
    .fn = ~str_replace(., "interassay", "Inter-assay component")
  ) %>%
  rename_with(
    .cols = matches("total"),
    .fn = ~str_replace(., "total", "Within-lab")
  ) %>%
  rename(
    c(
      # `Geometric mean (IU/mL)` = geo_mean,
    ` ` = panel_member, 
    `  ` = measurement,
    `Expected measurement` = expected_measurement
    )
  ) %>% 
  rename_with(
    .cols = matches("GSD"),
    .fn = ~str_replace(., "GSD", "Geometric SD")
  ) %>%
  rename_with(
    .cols = matches("GCV"),
    .fn = ~str_replace(., "GCV", "%GCV")
  ) %>%
  mutate(across(matches("CV|SD"), ~ replace_na(., 0))) %>%
  mutate(across(matches("CV"), ~ round(., 1))) %>%
  mutate(across(matches("SD"), ~ round(ifelse (. == 0, 0, exp(.)), 2))) %>%
  mutate(across(matches("mean"), ~ round(., 0))) %>%
  arrange(` `) %>%
  arrange(`  `) %>%
  flextable() %>%
  # span_header(sep = "_", i = 1, j = 2:8) %>%
  span_header(sep = "_") %>%
  vline(j = c(1, 2, 3, 4, 5, 7, 9)) %>%
  hline(part = "header", i = 1, j = 6:10) %>%
  hline(i = 4) %>%
  merge_h(part = "header", i = 1) %>%
  merge_v(j = 1) %>%
  align(part = "all", align = "center")

tab_3
```

# table 4 - rubella precision
```{r}
tab_4_df = read_excel("paper_audit.xlsx", sheet = "Validation; Precision rubella") %>%
  mutate(panel_member = `sample ID`) %>%
  group_by(panel_member) %>%
  mutate(
    n = n(), 
    ## ND50
    IU_50 = ND50 * 0.37,
    # IU_50 = ND50 * 0.370027752081406,
    logIU_50 = log(IU_50),
    ## ND80
    IU_80 = ND80 * 1.32,
    # IU_80 = ND80 * 1.32013201320132,
    logIU_80 = log(IU_80),
    # ND80_geomean = geo_mean(IU_80),
    ) %>%
  ungroup() %>%
  mutate(day = case_when(
    str_detect(plateID, "2023-May-01") ~ 2,
    str_detect(plateID, "2023-May-03") ~ 3
  ))

anova_tab_50 <- anovaVCA(logIU_50 ~ day,
                      Data = as.data.frame(tab_4_df),
                      by = "sample ID",
                      NegVC = TRUE)


tab_4_anova_50 <- lapply(names(anova_tab_50), function(sample_id) {
  vca_obj <- anova_tab_50[[sample_id]]
  tab = as.data.frame(vca_obj[["aov.tab"]])
  
  br_v = (tab[["MS"]][2] - tab[["MS"]][3]) / 2
  wr_v = tab[["MS"]][3]
  tot_v = br_v + wr_v
  
  br_sd = sqrt(br_v)
  wr_sd = sqrt(wr_v)
  tot_sd = sqrt(tot_v)
  
  br_cv = calculate_GCV(br_sd)
  wr_cv = calculate_GCV(wr_sd)
  tot_cv = calculate_GCV(tot_sd)
  
  gmeans = tab_4_df %>%
    group_by(panel_member) %>%
    reframe(panel_member = panel_member, ND50_geo_mean = geo_mean(IU_50)) %>%
    distinct()
  
  left_join(data.frame(
    measurement = "ND50",
    panel_member = str_split_i(sample_id, "sample ID.", 2),
      
    ND50_intraassay_GSD = wr_sd,
    ND50_intraassay_GCV = wr_cv,
    
    ND50_interassay_GSD = br_sd,
    ND50_interassay_GCV = br_cv,
    
    ND50_total_GSD = tot_sd,
    ND50_total_GCV = tot_cv
  ), gmeans, by = "panel_member")
})

anova_tab_80 <- anovaVCA(logIU_80 ~ day,
                      Data = as.data.frame(tab_4_df),
                      by = "sample ID",
                      NegVC = TRUE)

tab_4_anova_80 <- lapply(names(anova_tab_80), function(sample_id) {
  vca_obj <- anova_tab_80[[sample_id]]
  tab = as.data.frame(vca_obj[["aov.tab"]])
  
  br_v = (tab[["MS"]][2] - tab[["MS"]][3]) / 2
  wr_v = tab[["MS"]][3]
  tot_v = br_v + wr_v
  
  br_sd = sqrt(br_v)
  wr_sd = sqrt(wr_v)
  tot_sd = sqrt(tot_v)
  
  br_cv = calculate_GCV(br_sd)
  wr_cv = calculate_GCV(wr_sd)
  tot_cv = calculate_GCV(tot_sd)
  
  gmeans = tab_4_df %>%
    group_by(panel_member) %>%
    reframe(panel_member = panel_member, ND80_geo_mean = geo_mean(IU_80)) %>%
    distinct()
  
  left_join(data.frame(
    measurement = "ND80",
    panel_member = str_split_i(sample_id, "sample ID.", 2),
    
    ND80_intraassay_GSD = wr_sd,
    ND80_intraassay_GCV = wr_cv,
    
    ND80_interassay_GSD = br_sd,
    ND80_interassay_GCV = br_cv,
    
    ND80_total_GSD = tot_sd,
    ND80_total_GCV = tot_cv
  ) %>% mutate(
    expected_measurement = case_when(
      panel_member == "R-01" ~ 1379.5 * 0.37,
      panel_member == "R-02" ~ 396.8 * 0.37,
      panel_member == "R-08" ~ 2265.4 * 0.37,
      panel_member == "R-12" ~ 3219.4 * 0.37,
      panel_member == "R-14" ~ 3696.4 * 0.37,
    )), gmeans, by = "panel_member")
})

# anova_df = bind_rows(tab_4_anova_50, tab_4_anova_80)
tab_4_df = left_join(tab_4_df, bind_rows(tab_4_anova_50), by = "panel_member")
tab_4_df = left_join(tab_4_df, bind_rows(tab_4_anova_80), by = "panel_member")

tab_4 = tab_4_df %>%
  select(panel_member, n, ND50_geo_mean, ND50_total_GSD, ND50_total_GCV, ND80_geo_mean, ND80_total_GSD, ND80_total_GCV) %>%
  # mutate(expected_measurement = round(expected_measurement, 0)) %>%
  mutate(across(matches("CV"), ~ round(., 1))) %>%
  mutate(across(matches("SD"), ~ round(exp(.), 2))) %>%
  mutate(across(matches("mean"), ~ round(., 0))) %>%
  distinct() %>%
  arrange(panel_member) %>%
  rename(
    ` ` = panel_member
  ) %>% 
  rename_with(
    .cols = matches("geo_mean"), 
    .fn = ~str_replace(., "geo_mean", "Geometric mean")
  ) %>%
  rename_with(
    .cols = matches("GSD"),
    .fn = ~str_replace(., "_.*GSD", "_Geometric SD")
  ) %>%
  rename_with(
    .cols = matches("GCV"),
    .fn = ~str_replace(., "_.*GCV", "_%GCV (within lab)")
  ) %>%
  rename_with(
    .cols = matches("ND50"),
    .fn = ~str_replace(., "ND50", "IU/mL from ND50")
  ) %>%
  rename_with(
    .cols = matches("ND80"),
    .fn = ~str_replace(., "ND80", "IU/mL from ND80")
  ) %>%
  # rename(
    # `Expected measurement` = expected_measurement
  # ) %>%
  flextable() %>%
  span_header(sep = "_") %>%
  vline(part = "all", j = c(1, 2, 5, 8)) %>%
  # vline(part = "all", j = c(1, 2, 3, 6, 9)) %>%
  # hline(part = "header", i = 1, j = 3:8, border = fp_border_default(width = 2)) %>% 
  # vline(part = "all", j = 8) %>%
  merge_h(i = 1, part = "header") %>%
  # merge_v(j = 3:8) %>%
  align(part = "all", align = "center")

tab_4
```

# table 5 - RSV B precision
```{r}
tab_5_df = read_excel("paper_audit.xlsx", sheet = "Validation; RSV B") %>%
  filter(`sample ID` != "negative control") %>%
  mutate(panel_member = `sample ID`) %>%
  group_by(panel_member) %>%
  mutate(
    n = n(), 
    ## ND50
    IU_50 = ND50 * 0.91,
    # IU_50 = ND50 * 0.910875297,
    logIU_50 = log(IU_50),
    ) %>%
  mutate(expected_50 = case_when(
    panel_member == "2" ~ 1386,
    panel_member == "3" ~ 346,
    panel_member == "4" ~ 87,
    panel_member == "5" ~ 22
  )) %>%
  mutate(expected_80 = case_when(
    panel_member == "2" ~ 1263,
    panel_member == "3" ~ 315,
    panel_member == "4" ~ 79,
    panel_member == "5" ~ 20
  )) %>%
  mutate(day = case_when(
    str_detect(run_date, "2023-12-04") ~ 1,
    str_detect(run_date, "2023-12-05") ~ 2,
    str_detect(run_date, "2023-11-27") ~ 3
  ))

anova_tab_50 <- anovaVCA(logIU_50 ~ day,
                      Data = as.data.frame(tab_5_df),
                      by = "sample ID",
                      NegVC = TRUE)

tab_5_anova_50 <- lapply(names(anova_tab_50), function(sample_id) {
  vca_obj <- anova_tab_50[[sample_id]]
  tab = as.data.frame(vca_obj[["aov.tab"]])
  
  br_v = (tab[["MS"]][2] - tab[["MS"]][3]) / 2
  wr_v = tab[["MS"]][3]
  tot_v = br_v + wr_v
  
  br_sd = sqrt(br_v)
  wr_sd = sqrt(wr_v)
  tot_sd = sqrt(tot_v)
  
  br_cv = calculate_GCV(br_sd)
  wr_cv = calculate_GCV(wr_sd)
  tot_cv = calculate_GCV(tot_sd)
  
  gmeans = tab_5_df %>%
    group_by(panel_member) %>%
    reframe(panel_member = panel_member, mean = mean(IU_50), geo_mean = geo_mean(IU_50)) %>%
    distinct()
  
  left_join(data.frame(
    measurement = "ND50",
    panel_member = str_split_i(sample_id, "sample ID.", 2),
    
      
    intraassay_GSD = wr_sd,
    intraassay_GCV = wr_cv,
    
    interassay_GSD = br_sd,
    interassay_GCV = br_cv,
    
    total_GSD = tot_sd,
    total_GCV = tot_cv
  ), gmeans, by = "panel_member")
})

tab_5_df = left_join(tab_5_df, bind_rows(tab_5_anova_50), by = "panel_member")

tab_5 = tab_5_df %>%
  select(
    panel_member, 
    n, 
    geo_mean,
    # mean,
    matches(".*_GSD|.*_GCV")
  ) %>%
  distinct() %>%
  rename_with(
    .cols = matches("intraassay"),
    .fn = ~str_replace(., "intraassay", "Intra-assay component")
  ) %>%
  rename_with(
    .cols = matches("interassay"),
    .fn = ~str_replace(., "interassay", "Inter-assay component")
  ) %>%
  rename_with(
    .cols = matches("total"),
    .fn = ~str_replace(., "total", "Within-lab")
  ) %>%
  rename(
    c(`Geometric mean (IU/mL)` = geo_mean,
    ` ` = panel_member
    )
  ) %>%
  rename_with(
    .cols = matches("GSD"),
    .fn = ~str_replace(., "GSD", "Geometric SD")
  ) %>%
  rename_with(
    .cols = matches("GCV"),
    .fn = ~str_replace(., "GCV", "%GCV")
  ) %>%
  mutate(across(matches("CV|SD"), ~ replace_na(., 0))) %>%
  mutate(across(matches("CV"), ~ round(., 1))) %>%
  mutate(across(matches("SD"), ~ round(ifelse (. == 0, 0, exp(.)), 2))) %>%
  mutate(across(matches("mean"), ~ round(., 0))) %>%
  distinct() %>%
  arrange(` `) %>%
  flextable() %>%
  span_header(sep = "_") %>% 
  vline(part = "all", j = c(1, 2, 3, 5, 7)) %>%
  # hline(part = "header", i = 1, j = 3:8, border = fp_border_default(width = 2)) %>% 
  # vline(part = "all", j = 8) %>%
  merge_h(i = 1, part = "header") %>%
  # merge_v(j = 3:8) %>%
  align(part = "all", align = "center") %>%
  autofit()

tab_5
```

```{r}
b_df = read_excel("neut analysis EP/B_analysis.xlsx") %>%
  filter(virus %in% c("B150", "B200"))


tab_5_b150 = b_df %>%
  filter(virus == "B150") %>%
  filter(!str_detect(sample_base, "negative")) %>%
  select(sample_base, ND50, virus) %>%
  group_by(sample_base) %>%
  summarise(
    n = n(),
    geo_mean = round(geo_mean(ND50 * 0.91), 0),
    geo_mean_unrounded_cf = round(geo_mean(ND50 * 0.910875297), 0),
    mean = round(mean(ND50 * 0.91), 0),
    mean_unrounded_cf = round(mean(ND50 * 0.910875297), 0),
    virus = virus) %>%
  distinct() %>%
  ungroup() %>%
  flextable() %>%
  align(part = "all", align = "center") %>%
  add_footer_lines("Using all 6 ND50 values sourced from B150")

tab_5_b150

tab_5_b200 = b_df %>%
  filter(virus == "B200") %>%
  filter(!str_detect(sample_base, "negative")) %>%
  select(sample_base, ND50, virus) %>%
  group_by(sample_base) %>%
  summarise(
    n = n(),
    geo_mean = round(geo_mean(ND50 * 0.91), 0),
    geo_mean_unrounded_cf = round(geo_mean(ND50 * 0.910875297), 0),
    mean = round(mean(ND50 * 0.91), 0),
    mean_unrounded_cf = round(mean(ND50 * 0.910875297), 0),
    virus = virus) %>%
  distinct() %>%
  ungroup() %>%
  flextable() %>%
  align(part = "all", align = "center") %>%
  add_footer_lines("Using all 6 ND50 values sourced from B200")

tab_5_b200

tab_5_combined = b_df %>%
  filter(!str_detect(sample_base, "negative")) %>%
  select(sample_base, ND50, virus) %>%
  group_by(sample_base) %>%
  summarise(
    n = n(),
    geo_mean = round(geo_mean(ND50 * 0.91), 0),
    geo_mean_unrounded_cf = round(geo_mean(ND50 * 0.910875297), 0),
    mean = round(mean(ND50 * 0.91), 0),
    mean_unrounded_cf = round(mean(ND50 * 0.910875297), 0)
  ) %>%
  distinct() %>%
  ungroup() %>%
  flextable() %>%
  align(part = "all", align = "center") %>%
  add_footer_lines("Using all 12 ND50 values sourced from B150 and B200")

tab_5_combined
```

## regenerate Table S6
```{r}
all_df = read_excel("paper_audit.xlsx", sheet = "Results") %>%
  filter(paper == "yes")

table_s6 = all_df %>%
  mutate(group = case_when(
    sample_type2 == "timepoint" ~ "HSV serology remnant specimens",
    sample_type2 == "future RSV-infected" ~ "Future RSV PCR+",
    sample_type2 == "RSV+, immunocompetent" ~ "RSC PCR+ immunocompetent",
    sample_type2 == "RSV+, immunosuppressed" ~ "RSV PCR+ immunosuppressed",
    sample_type2 == "Influenza+, RSV-" ~ "Influenza PCR+, RSV PCR-",
    str_detect(`sample ID`, "^R-") ~ "Rubella serology remnant specimens",
    .default= "HSV serology remnant specimens"
  ),
  `Age group` = case_when(
    pat_age < 18 ~ "below 18",
    pat_age >= 18 & pat_age <= 65 ~ "18-65",
    pat_age > 65 ~ "65+",
  ),
  `Immunosuppression (yes/no)` = case_when(
    sample_type2 == "RSV+, immunocompetent" ~ "N",
    sample_type2 == "RSV+, immunosuppressed" ~ "Y",
    .default = "Unknown"
  ),
  Individual = ifelse(group == "Rubella serology remnant specimens", `sample ID`, dummy),
  Sex = PatSex,
  Age = pat_age,
  `Hospital source` = performing_lab_locs,
  `RSV Neutralizing Titer (IU/mL)` = ifelse(group == "Rubella serology remnant specimens", round(`IU/mL RSV A2`, 0), round(ND50 * 0.37, 0)), 
  `Rubella Neutralizing Titer (IU/mL)` = round(`Rubella IU/mL`, 0),
  ) %>%
  select(
    Individual, 
    `RSV Neutralizing Titer (IU/mL)`, 
    `Rubella Neutralizing Titer (IU/mL)`,
    `Immunosuppression (yes/no)`,
    `Hospital source`,
    `Sex`,
    `Age`,
    `Age group`,
    `group`
  ) %>%
  distinct() %>%
  arrange(desc(`RSV Neutralizing Titer (IU/mL)`)) %>%
  arrange(group)

table_s6 %>%
  flextable() %>%
  align(part = "all", align = "center") %>%
  autofit()

write.xlsx(table_s6, "paper_audit_figs/Table S6_raw.xlsx")

```